---
title: "Travelpac Analysis Project"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Overseas Travel and Tourism: An Analysis ###

## The data ##
The Travelpac dataset, found on the Office for National Statistics website here (https://www.ons.gov.uk/peoplepopulationandcommunity/leisureandtourism/datasets/travelpac), contains estimates of international visits with details on traveller age and sex, trip purpose, length, and spending from the International Passenger Survey (IPS). 


## Our Aims ##

We wish to generate insights on tourism trends and habits, which can be used in a multitude of applications. We will pretend to work for a fictional travel agent. We wish to maximise our profit on our new luxury travel package, and need to work out who to advertise this to.

### Set up ###

# Packages #
We load our packages. The tidyverse library contains the dplyr package, which will be useful for data handling, and the ggplot2 package, which we will use for visualisations.

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
```

# Data #

Let us load in our raw data.

```{r}
travel_raw=read_excel('travelpac_raw/Travelpac 2023.xlsx', 
                      sheet=2)
```

### Exploratory Analysis ###

# Summary #

Let's take a quick look at the data structure.

```{r}
summary(travel_raw)
```

We have some categorical variables, which we can use to stratify the data, such as sex, age group and travel purpose, and we have some numeric results, from which we can generate insights.

# Cleaning #

We notice from looking at the dataset, that we have some NAs in the Sex variable. It is not clear how best to impute these, so we will just exclude these rows from the analysis.

```{r}
travel_cleaned <- travel_raw %>% drop_na(Sex)
```

We also have 'D/K' terms appearing in the age group column. Let's remove these.

```{r}
travel_cleaned2 <- travel_cleaned[travel_cleaned$Age != 'D/K',]
```

